---
title: "IPL Data Analysis"
author: "Srinivas Shekar 01FB15ECS308"
date: ' By 17 August 2017'
output:
  word_document: default
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Loading Data:

```{r}
data<-read.csv("deliveries.csv",sep=",")
result1<-read.csv('matches.csv',sep=',')
```

Loading required libraries:
```{r message=FALSE}
library(magrittr)
library(dplyr)
library(ggplot2)
library(gdata)

```



##Question 1:Analysis on IPL Final 2016


1.Loading required data :

```{r}
d<-subset(data,match_id==577)
data_srh<-subset(d,batting_team=='Sunrisers Hyderabad')
data_rcb<-subset(d,batting_team=='Royal Challengers Bangalore')
```

2.Target set by SRH :

```{r}
s<-sum(data_srh$total_runs)
print(s)
```

3."Beat the Run Rate Boys!!"- V Kohli

```{r}
run_rate<-s/max(data_srh$over)
print(run_rate)
```

4.RCB's batting analyis :

```{r}
df_plot <-data_rcb %>% group_by(over)%>% summarise(runs=sum(total_runs)) %>% arrange(over)

df_plot %>% ggplot(aes(over,runs))+geom_histogram(stat="identity",color='blue',fill='cyan')+xlab('Overs')+
  ylab('Total runs scored')+ggtitle('Total runs scored by RCB in each over')
```

##Question 2:Analysis on team's batting performance

1.Hungry for runs???Here is the total runs scored by teams in all seasons

```{r}
d_7<-subset(data,(match_id %in% c(7,27,67,171,414)) & is_super_over==0)
d_7 %>%  group_by(inning,match_id,batting_team) %>% summarise(total_runs=sum(total_runs)) %>% arrange(match_id)
```

2.Record breakers!!

```{r message=FALSE}

result2<-subset(result1,(result %in% c('normal')) &(dl_applied==0))
result2 <- rename.vars(result2, from = "id", to = "match_id")
s1<-subset(data,match_id %in% result2$match_id)

q<-s1 %>%  group_by(match_id,inning,batting_team) %>% summarise(total_runs=sum(total_runs)) %>% arrange(match_id)

max_team<-subset(q,max(q$total_runs)==total_runs)
max_team
```

3.Oops!! Record breakers or Record ....

```{r}
min_team<-subset(q,min(q$total_runs)==total_runs)
min_team
```

##Question 3: IPL Auction:Who is the effective player??Lets look at marquee players

```{r}
getmode<-function(v){
  uniqv<-unique(v)
  uniqv[which.max(tabulate(match(v,uniqv)))]
}
```

1.MS Dhoni's IPL career:

```{r}
msd<-subset(data,batsman=='MS Dhoni')
msd<-msd %>% group_by(match_id,batsman) %>% summarise(total_runs=sum(total_runs)) %>% arrange(match_id)
with(msd,sum(total_runs))
with(msd,summary(total_runs))
with(msd,getmode(total_runs))
with(msd,IQR(total_runs))
with(msd,sd(total_runs))
```

2.V Kohli's IPL career:

```{r}
kohli<-subset(data,batsman=='V Kohli')
kohli<-kohli %>% group_by(match_id,batsman) %>% summarise(total_runs=sum(total_runs)) %>% arrange(match_id)
with(kohli,summary(total_runs))
with(kohli,sum(total_runs))
with(kohli,getmode(total_runs))
with(kohli,IQR(total_runs))
with(kohli,sd(total_runs))
```

3.ABD Villiers's IPL career:

```{r}
abd<-subset(data,batsman=='AB de Villiers')
abd<-abd %>% group_by(match_id,batsman) %>% summarise(total_runs=sum(total_runs)) %>% arrange(match_id)
with(abd,summary(total_runs))
with(abd,sum(total_runs))
with(abd,getmode(total_runs))
with(abd,IQR(total_runs))
with(abd,sd(total_runs))
```

4.Jadeja's IPL career:

```{r}
jaddu<-subset(data,batsman=='RA Jadeja')
jaddu<-jaddu %>% group_by(match_id,batsman) %>% summarise(total_runs=sum(total_runs)) %>% arrange(match_id)
with(jaddu,summary(total_runs))
with(jaddu,sum(total_runs))
with(jaddu,getmode(total_runs))
with(jaddu,IQR(total_runs))
with(jaddu,sd(total_runs))
```

5.Suresh Raina's IPL career:

```{r}
raina<-subset(data,batsman=='SK Raina')
raina<-raina %>% group_by(match_id,batsman) %>% summarise(total_runs=sum(total_runs)) %>% arrange(match_id)
with(raina,summary(total_runs))
with(raina,sum(total_runs))
with(raina,getmode(total_runs))
with(raina,IQR(total_runs))
with(raina,sd(total_runs))
```

##Question 4: MSD v/s Kohli

1.Comparitive Boxplots:
```{r}


boxplot(msd$total_runs , kohli$total_runs,col = terrain.colors(4))
legend("topleft",title="Batsman",c("MS Dhoni","V Kohli"),fill=terrain.colors(4),horiz=TRUE)
```

2.90th Percentile

```{r}
quantile(msd$total_runs,prob=0.9)
quantile(kohli$total_runs,prob=0.9)
```

3. (a) Top performances by Kohli

```{r}

kohli<-kohli %>% arrange(total_runs)
tail(kohli)
```

(b) Top performances by MSD

```{r}

msd<-msd %>% arrange(total_runs)
tail(msd)
```

4.(a) Worst performances by Kohli

```{r}
head(kohli,8)
```

(b) Worst performances by MSD
```{r}
head(msd,8)
```

##Question 5: Is Suresh Raina a key player in winning matches for Gujarat Lions?? 

1.Loading Gujarat Lion's Data...

```{r}
gl<-subset(data,batting_team=='Gujarat Lions')
gl<-gl %>% group_by(match_id) %>% summarise(total_runs=sum(total_runs)) %>% arrange(match_id)
```

2.Loading Raina's data...
```{r}
raina_gl<-subset(data,batting_team=='Gujarat Lions' & batsman=='SK Raina')
raina_gl<-raina_gl %>% group_by(match_id) %>% summarise(total_runs=sum(total_runs)) %>% arrange(match_id)
```

3.Looking at Suresh Raina's impact in the win percentage of the team

```{r}
raina_vs_gl<-merge(raina_gl,gl,by='match_id')
raina_vs_gl$impact<-round((raina_vs_gl[2]/raina_vs_gl[3])*100,2)
raina_vs_gl



win<-subset(result1,winner=="Gujarat Lions")
win1<-subset(raina_vs_gl,match_id %in% win$id)

colnames(win1)<-c("Match id","Raina's score","Team Score","Impact of Raina on the team")
win1 


```

